\chapter*{Acknowledgments}

I would like to thank Dr. Goasguen for bringing me in to the Cyber Infrastructure
Research Group. I have learned a great deal about distributed computing and XMPP
thanks to his guidance and motivation. I am also grateful for my many discussions
with Dr. Stevenson over the last few years discussing programming languages and
their histories, and just life in general.

Mike Murphy has my gratitude for his help trying to teach me the proper methods
for academic writing, the many design discussions we had, and for being a good
friend. I want to also thank Michael Fenn for his help and experience with using
the Palmetto cluster and how to work with KVM.

I would like to acknowledge Jerome Lauret of Brookhaven National Laboratory and
Matthew Walker of MIT for their help and feedback in running experiments with
the STAR VO, Xiaoshu Wang of the Renaissance Computing Institute (RENCI) for his
feedback and bug reporting, and David Wolinksy of the University of Florida for
his help in conducting IPOP scalability tests.

Nathan Fritz of \&yet and author of the Python XMPP library SleekXMPP, featured
in O'Reilly's \textit{XMPP: The Definitive Guide} deserves a great many thanks.
He has trusted me enough not only to grant me commit access to the main source
repository for the SleekXMPP library, but to also grant me co-authorship status as
well. I hope that trust has not been misplaced; as a direct result of the work
presented here, SleekXMPP has been expanded and refined from what Nathan had
called a ``hobbyist project'' to a production ready 1.0 release. Features developed
as part of Kestrel that were generic in nature have made their way into the main
SleekXMPP library.
